Rare Pathing

### 2025-04-09 notebooks/rare_path.qmd
Update rare path for example graph diagrmans and analysis

## Direct rare pathing
```{python}
import igraph as ig
import numpy as np
import network_simulation as netsim

@@

R = ig.Graph(
    n=4, 
    edges=[
        (0, 1),
        (1, 2),
        (2, 0),
        (0, 3),
    ], 
    directed=True, 
    edge_attrs={'weight':[
        0,
        0,
        0,
        100,
    ]}
)
```

@@

```{python}
N_SIMS = 1000

histories = []
for it_ in range(N_SIMS):
    snapshot_sim.reset(soft=True)
    snapshot_sim.simulate(30, nostop=True)
    histories.append(np.array(snapshot_sim.history))
```

>>>>

## Direct rare pathing
```{python}
import igraph as ig
import numpy as np
from echidna import network_simulation as netsim

@@

R = ig.Graph(
    n=4, 
    edges=[
        # (0, 1),
        # (1, 2),
        # (2, 0),
        (0, 3),
    ], 
    directed=True, 
    edge_attrs={'weight':[
        # 0,
        # 0,
        # 0,
        100,
    ]}
)
```

```{python}
#| label: fig-graphs
fig, axs = plt.subplots(ncols=2)
axz = axs.flatten()

ig.plot(
    G,
    target=axz[0],
    edge_label=G.es['weight'],
    layout='circle',
)

ig.plot(
    R,
    target=axz[1],
    edge_label=R.es['weight'],
    layout='circle',
)

axz[0].set_xlabel('regular')
axz[0].spines[:].set_visible(True)
axz[1].set_xlabel('rare')
axz[1].spines[:].set_visible(True)
```

@@

run and analyse the snapshot stuff
```{python}
N_SIMS = 1000

snapshot_histories = []
for it_ in range(N_SIMS):
    snapshot_sim.reset(soft=True)
    snapshot_sim.simulate(30, nostop=True)
    snapshot_histories.append(snapshot_sim.history)
```

```{python}
timez = [snapshot_sim.ts[x] 
         for hist in snapshot_histories 
         for x in np.flatnonzero(hist[-1,:])[:1]
        ]

plt.hist(timez)
```

run and analyse static
```{python}
static_histories = []
for it_ in range(N_SIMS):
    static_sim.reset(soft=True)
    static_sim.simulate(30, nostop=True)
    static_histories.append(static_sim.history)
```

```{python}
static_timez = [static_sim.ts[x] 
         for hist in static_histories 
         for x in np.flatnonzero(hist[-1,:])[:1]
        ]
```
```{python}
#| label: fig-cumhist
plt.xlabel('hitting time')
# plt.ylabel('count')
# plt.hist(static_timez, histtype='step', bins=31, label='static')
# plt.hist(timez, histtype='step', bins=4, label='snapshot')
plt.ylabel('cumulative count')
plt.hist(static_timez, cumulative=True, histtype='step', bins=31, label='static')
plt.hist(timez, cumulative=True, histtype='step', bins=4, label='snapshot')
plt.legend()
# plt.yscale('log')
```
